A Middleware for Symmetric Peer-to-Peer Applications

### 2023-03-10 tpds-23/tpds-23.tex
...

 6. ***How does the middleware handle consistency when advancing or rollbacking in  time to synchronize?***
About #6, this is a valid criticism and we should include a brief discussion about consistency models and position our work on this spectrum.
The facts that peers share a global clock and go through bit-identical steps may characterize strict consistency.
However, since the global clocks may diverge temporarily, we must rely on rollbacks.
The key to preserve consistency in the presence of rollbacks is to restrict our programming API to deterministic and stateless calls only.
This way, it is not possible to diverge from other peers, for example, by re-executing stateful file operations multiple times (or at different times).

&gt; A apresentação da solução entra em um nível de detalhamento muito baixo...
  ...na decisão da topologia utilizada na simulação não fica muito claro se  essa topologia representa algum cenário específico.

Sobre a topologia escolhida, optamos por misturar 3 arranjos abstratos:

1. linha contínua
2. ciclo fracamente conectado
3. ciclo totalmente conectado

Entendemos que casos concretos que sejam subconjuntos dos arranjos acima terão desempenho igual ou superior ao exibido na Figura 7.
Por exemplo, uma linha contínua de até 15 nós é um subconjunto da topologia escolhida, assim como dois ciclos conectados separados por um nó.
Assim, pretendemos atender a um número maior de cenários, mesmo que isso dependa de uma interpretação mais apurada por parte dos interessados.
De qualquer maneira, a escolha é arbitrária e não representa um cenário concreto.

&gt; Figures 1 and 2 create an expectation of approaches comparison; I suggest writing in the motivation why there will be no clear comparison in the
  experimental results.
We will be more explicit about this in Sections 2/4, since it is not possible to reproduce our experiment in centralized solutions (the topology is inherently peer-to-peer).
We only discuss the performance of related works based on their claims and also on theoretical upper limits."

Os autores mencionam que em cenário com latência de 50ms e 25 eventos por minuto, os pares precisaram retroceder 3\% do tempo. Eles devem deixar claro porque isso é uma vantagem e qual a média de período de tempo aceitável para retrocesso.
- Nos trabalhos relacionados, apenas propostas centralizadas são mencionadas e comparadas. Não há propostas descentralizadas para comparação também? Sugiro expandir essa seção.
- O design experimental precisa ser melhor descrito. Apenas o emulador das latências é descrito (NetEm). Como foi realizada (ou qual ferramenta foi usada) a simulação dos experimentos?
- Apenas aplicações com deadline mais curtos adequam-se melhor ao middleware.  Os autores precisam explicar no texto os resultados dos experimentos com maior latência também (100ms, 250ms e 500ms).
- Os autores precisam deixar claro o benefício de usar uma solução descentralizada desde que o seu desempenho é menor que as soluções centralizadas.

@@

In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications}, which do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to a previous state and reapply events in order and
in time.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate $\Delta$ so that they can reach other peers in time to be applied in
sync.
%We first assume a predetermined delay for the sake of simplicity.
Nevertheless, a peer ahead of time may receive an event that should have been
applied in the past. %, even considering its $\Delta$.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate the whole
application execution and event occurrences from the beginning.
As an optimization, the middleware takes periodic snapshots locally to avoid
full simulation.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because applications are launched locally at different times, and
because internal clocks may diverge over time (e.g., from rollbacks or timer
polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
To avoid abrupt visual changes, the middleware makes smooth time transitions
that take up to $1s$ to complete.
%
\item \textbf{Peer Churn:}
Considering that we target dynamic peer-to-peer networks, it is important that
the middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware needs to ensure that it
receives from other peer all events that ever happened, in order.
This is trivial since all peers already need to hold the full event history (as
described in item (a)).
% including their absolute timestamps in the shared timeline.
\end{enumerate}

@@

This section provides an overview of the middleware implementation to answer
how it ensures the desired properties in concrete terms as follows:
    (a) peers meet event deadlines either travelling back and forth or in real
        time (Section \ref{sec.tml.middleware.loop}: \linx{12--16} or
        \linx{18--22});
    (b) peers synchronize their clocks with periodic \code{SYNC} packets that
        fast forward late peers (Section \ref{sec.tml.middleware.loop}:
        \linx{7--10}); and
    (c) peers can join the network at any time and still catch up in time by
        receiving the full event history
        (Section~\ref{sec.tml.middleware.events}).

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
%
We run the application main loop at $50~FPS$ ($20ms$ per frame) and calculate
the middleware overhead at each iteration.
We found a negligible overhead of $9us$ in the average, which corresponds to
less than $1/2000$ of the CPU time.
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency ($5-500ms$ between each peer),
    (ii)  the rate of events ($5-200~evt/sec$), and
    (iii) the deadline $\Delta$ multiplier ($1-5x$).
%
Figure~\ref{fig.mult} shows the $\Delta$s considering the network latency and
chosen multiplier.
For instance, a $50ms$ latency with a $2x$ multiplier results in
$\Delta=500ms$, i.e., each event is triggered with this deadline delay to
compensate the given latency and fixed average of 5 hops
($5 \times 50 \times 2 = 500$).

@@

local-first
Apesar dos rollbacks quando a rede se reconecta serem inevitáveis, os seus efeitos podem ser imperceptíveis para os usuários finais (e.g., quando há pouca
atividade ou dependência entre eventos).

O principal benefício é que a aplicação continua funcionando mesmo com a
presença de partições na rede, o que inclui a queda do (que seria o) servidor
em uma solução centralizada.

Levando essa possibilidade ao extremo, um peer ainda poderia funcionar
normalmente mesmo que totalmente offline por um determinado período (como é
comum em pesquisas recentes em &quot;local-first software&quot;).

>>>>

\begin{comment}
Os autores mencionam que em cenário com latência de 50ms e 25 eventos por
minuto, os pares precisaram retroceder 3\% do tempo.
Eles devem deixar claro porque isso é uma vantagem e qual a média de período de
tempo aceitável para retrocesso.

- O design experimental precisa ser melhor descrito.
Apenas o emulador das latências é descrito (NetEm).
Como foi realizada (ou qual ferramenta foi usada) a simulação dos experimentos?

- Apenas aplicações com deadline mais curtos adequam-se melhor ao middleware.
Os autores precisam explicar no texto os resultados dos experimentos com maior
latência também (100ms, 250ms e 500ms).
\end{comment}

@@

In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications}, which do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a distributed time
machine can rollback peers to previous states and reapply the events in order,
such that a single consistent timeline is established across the network.
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.

@@

\subsection{Conflict-free Replicated Data Types}

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to provide a consistent timeline and preserve the symmetric
behavior across peers.
As described in the Introduction, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate $\Delta$ so that they can reach other peers in time to be applied in
sync.
%We first assume a predetermined delay for the sake of simplicity.
Nevertheless, a peer ahead of time may receive an event that should have been
applied in the past. %, even considering its $\Delta$.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
As we detail further, rolling back a peer requires to simulate the whole
application execution and event occurrences from the beginning.
As an optimization, the middleware takes periodic snapshots locally to avoid
full simulation.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because applications are launched locally at different times, and
because internal clocks may diverge over time (e.g., from rollbacks or timer
polling inaccuracies).
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
To avoid abrupt visual changes, the middleware makes smooth time transitions
that take up to $1s$ to complete.
%
\item \textbf{Peer Churn:}
Considering that we target dynamic peer-to-peer networks, it is important that
the middleware can handle arrival and departure of peers seamlessly.
Note that high churn may lead to intermittent network partitions.
Given our unstructured approach, nothing needs to be done when a peer leaves
the network.
However, when a peer joins the network, the middleware needs to ensure that it
receives from other peer all events that ever happened, in order.
This is trivial since all peers already need to hold the full event history (as
described in item (a)).
% including their absolute timestamps in the shared timeline.
\end{enumerate}

@@

This section provided an overview of the middleware implementation to answer
how it ensures the desired properties in concrete terms as follows:
    (a) peers meet event deadlines either travelling back and forth or in real
        time (Section \ref{sec.tml.middleware.loop}: \linx{12--16} or
        \linx{18--22});
    (b) peers synchronize their clocks with periodic \code{SYNC} packets that
        fast forward late peers (Section \ref{sec.tml.middleware.loop}:
        \linx{7--10}); and
    (c) peers can join the network at any time and still catch up in time by
        receiving the full event history
        (Section~\ref{sec.tml.middleware.events}).

The fact that peers share a global timeline of event, and that they go through
bit-identical steps ensures consistency in the network.
However, since the clocks may diverge temporarily, we must rely on rollbacks.
Therefore, the key to preserve consistency in the presence of rollbacks is to
restrict our programming API to deterministic and stateless calls only.
This way, it becomes impossible to diverge from other peers by, for example,
re-executing stateful file operations multiple times (or at different times).

@@

We perform simulations with 21 peers using the mixed topology of
Figure~\ref{fig.topo}, with peers organized in straight lines and cycles.
The topology has an average of 5 hops between peers (e.g., \code{4-16} are 7
hops away).
We opted to mix three abstract arrangements:
    (i) straight lines, and (ii) weakly and (iii) strongly connected cycles.
Even though this topology is arbitrary, we can expect that concrete scenarios
with subsets of these arrangements will have at least the same performance of
our experiments.
For instance, a straight line of 15 peers is a subset of our topology, as well
as two cycles separated by a single peer.

We run the application main loop at $50~FPS$ ($20ms$ per frame) and calculate
the middleware overhead at each iteration.
We found a negligible overhead of $9us$ in the average, which corresponds to
less than $1/2000$ of the CPU time.
%
In order to have an absolute clock, we use a single machine to simulate all
peers, allowing us to merge their logs into a single file with a comparable
shared timeline.
%
We use \emph{NetEm}~\cite{netem} to simulate the network latency with a normal
distribution.
% (e.g., \code{tc qdisc add dev lo root netem delay 200ms 40ms distribution normal}).
%
We vary
    (i)   the network latency ($5-500ms$ between each peer),
    (ii)  the rate of events ($5-200~evt/sec$), and
    (iii) the deadline $\Delta$ multiplier ($1-5x$).
%
Figure~\ref{fig.mult} shows the $\Delta$s considering the network latency and
chosen multiplier.
For instance, a $50ms$ latency with a $2x$ multiplier results in
$\Delta=500ms$, i.e., each event is triggered with this deadline delay to
compensate the given latency and fixed average of 5 hops
($5 \times 50 \times 2 = 500$).

@@
